---
title: "Assignment 01"
author: "EDDA-Group43"
date: "2021/2/19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1. Birthweights
```{r}
data = read.table(file="birthweight.txt", header=TRUE)
qqnorm(data$birthweight)
data_mean = mean(data$birthweight)
error = qt(0.95, df=length(data$birthweight)) * sd(data$birthweight) / sqrt(length(data$birthweight))
upper_bound = data_mean + error
lower_bound = data_mean - error
```


```{r}
t.test(data$birthweight, conf.level=0.9)
```


```{r}
t.test(data$birthweight, mu=2800, alternative="greater")
```








# Exercise 5. Chick weights
## a.
```{r}
data("chickwts")
meatmeal = chickwts$weight[chickwts$feed == "meatmeal"]
sunflower = chickwts$weight[chickwts$feed == "sunflower"]
boxplot(meatmeal, sunflower, names=c("meatmeal", "sunflower"))
```

```{r}
t.test(meatmeal, sunflower, paired=FALSE)
```
Since the Welch Two Sample t-test has $p$-value=0.04441 < 0.05, it is the case to reject $H_0$. The conclusion should be there exist certain difference between the meatmeal and sunflower. Additionally, the data are not paired, because if we set "paired=TRUE", we would get an error message- "not all arguments have the same length".

```{r}
wilcox.test(meatmeal, sunflower)
```
Mann-Whitney test has a $p$-value = 0.06882 > 0.05, so the conclusion is that there is no such significant difference between two groups.

```{r}
ks.test(meatmeal,sunflower)
```
Kolmogorov-Smirnov test has a $p$-value = 0.1085 < 0.05. From this result we can conclude that the two populations are extremely unsymmetrical in shapes.

## b.
```{r}
chickaov=lm(weight ~ feed, data=chickwts)
anova(chickaov)
```
By conducting the one-way ANOVA, it is easy to see that the $p$-value = 5.936e-10 < 0.05, which tells that there exist at least such a type has an extremely different average weight. 
```{r}
library(ggplot2)
library(ggthemes)
library(dplyr)
cdata = tbl_df(chickwts)
cdata.wt.feed = cdata %>%
  ggplot(aes(x=feed, y=weight)) +
  geom_count(color="red") +
  labs(title="Relationship Between Chick Weight & Type of Feed",
       x="Feed Type", y="Weight in Grams") + 
  theme_calc()
cdata.wt.feed
```
The above plot shows that the casein and sunflower are the best feed supplement.

## c.
```{r}
caov = aov(weight~feed, data=chickwts)
plot(caov)
```
These plots shows that the ANOVA model assumptions is in line with expectations. In the case like this, we can conclude that values have equal variance. 
```{r}
summary(caov)
```
Additionally, since the $p$-value < 0.05, it also shows the normality of the values.

## d.
```{r}
kruskal.test(weight ~ feed, data=chickwts) 
```
Kruskal-Wallis test has a $p$-value = 5.113e-07 < 0.05, which means there exist such a feed influenced the weight. And the one-way ANOVA also shows the same conclusion. On the other hand, Kruskal-Wallis test is based on ranks instead of normality as ANOVA. It means that the ANOVA tested the normality of values from means, but Kruskal-Wallis tested on comparison of the ranks of the means.